ISODS: An ISODATA-Based Initial Centroid Algorithm

### 2021-01-08 paper/paper.tex
add references

\lettrine[nindent=0em,lines=3]{R} emote sensing technology is used when
the data collection takes place at a remote distance from the subject matter,
such as the study of daily weather and long-term climate change, land cover
monitoring, ecosystem modeling of vegetation, and so on. Multispectral imaging
technology is one of the instruments for remote sensing used since the late
1960s. However, multispectral imaging systems use a low number of spectral bands
(three to six spectral bands in different systems). Increasing spectral
resolution in the hyperspectral images (HSI) with several hundred spectral bands
of the observational scene in a single acquisition makes it possible to acquire
detailed examination of observational scenes. The ground-based version of
hyperspectral images has gained immense interest. It has been applied to
numerous electronic imaging applications, including food inspection, forensic
science, medical surgery and diagnosis, and military applications. Because of
the difficulty of obtaining prior knowledge in remote sensing studies,
unsupervised classification and cluster analysis have great importance compared
to supervised classification.

Unsupervised learning is a machine learning technique used when the learning
algorithm has no pre-existing labels and looks for patterns in the data set with
minimum human supervision and without external guidance. In contrast to
supervised learning, where a training set of inputs and observations are given
to the method, unsupervised learning uses only the inputs and should find an
interesting pattern in the data set. Two main approaches are used in
unsupervised learning, named principal component and cluster analysis. In the
clustering method, unlabeled and unsorted data points are grouped and segmented
according to similarities and differences to find structure in the inputs, group
it into meaningful clusters, and understand the data set better. There are
different algorithms for Clustering, such as K-Means, fuzzy C-Means, ISODATA,
and so on. In the K-Means algorithm, the goal is to minimize the sum of the
squared distances points have in the data set from its center. The problem with
the K-Means algorithm is that it requires a priori knowledge about the number of
classes in the data set, and choosing the cluster centers randomly may not
result in optimal partitioning of the data set.

One of the most popular clustering algorithms is called ISODATA clustering,
based on the K-means algorithm with some variations. The ISODATA method's
preference in comparison with the K-means algorithm is that the ISODATA
algorithm uses splitting and merging forms for clusters. The ISODATA method
allows starting from arbitrary cluster centers and obtains better partitioning
results than the K-Means algorithm, but it is not entirely resilient to initial
cluster centers. In this paper, a novel method using the ISODATA approach of
splitting clusters to select high-quality initial cluster centers is proposed.
This new method decreases the number of iterations and improves the overall
performance of the algorithm.

The paper is organized as follows: Section 2 introduces K-Means and ISODATA
clustering algorithms. The new proposed algorithm is explained in Section 3. We
present the experimental results in Section 4. Finally, Section 5 contains
concluding remarks.

@@

This method began as a modified ISODATA process. However, the modification was
unique in that there is a clear distinction from the proposed change and the
beginning of the original process. Although more complex than other methods of
generating initial cluster centers, this approach guarantees variance between
clusters. As a result, ISODS can be used to generate clusters for other
clustering algorithms, such as K-Means. In the next section, ISODS is applied to
both ISODATA and K-Means on some benchmark UCI datasets [CITE] and compared with
Random Centers and K-Means++ Centers.

@@

Due to the response of the ISODATA algorithm to the discarding, merging, and
splitting parameters, these parameters should be carefully selected. Since the
goal of this research is to explore the impact of initial cluster centers on
ISODATA, we use a relatively simple approach for generating ISODATA parameters.
Based on the available training data with included labels, each class is
summarized with the mean and standard deviation of that class, as well as the
number of members in each class. To determine $\theta n$, the minimum number of
cluster members used for discarding, we find the class with the fewest members
and use the number of members in that class as $\theta n$. For $\theta e$, the
maximum standard deviation allowed for a cluster, we follow a similar approach
and find the maximum standard deviation within a class. Finally, for $\theta c$,
we calculate the distance between the means of each class, and use the minimum
distance as $\theta c$.

@@

A statement requiring citation \cite{Figueredo:2009dg}.
\blindtext % Dummy text

@@

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
Figueredo, A.~J. and Wolf, P. S.~A. (2009).
\newblock Assortative pairing and life history strategy - a cross-cultural
  study.
\newblock {\em Human Nature}, 20:317--330.

\end{thebibliography}

>>>>

\lettrine[nindent=0em,lines=3]{R} emote sensing technology is used when the data
collection takes place at a remote distance from the subject matter, such as the
study of daily weather and long-term climate change, land cover monitoring,
ecosystem modeling of vegetation, and so on \cite{Jensen1996,Mallapragada2018}.
Multispectral imaging technology is one of the instruments for remote sensing
used since the late 1960s \cite{Khan2018,Landgrebe1999}. However, multispectral
imaging systems use a low number of spectral bands (three to six spectral bands
in different systems). Increasing spectral resolution in the hyperspectral
images (HSI) with several hundred spectral bands of the observational scene in a
single acquisition makes it possible to acquire detailed examination of
observational scenes \cite{Khan2018}. The ground-based version of hyperspectral
images has gained immense interest. It has been applied to numerous electronic
imaging applications, including food inspection, forensic science, medical
surgery and diagnosis, and military applications
\cite{Govender2009,Adam2009,Fischer2006,Carrasco2003}. Because of the difficulty
of obtaining prior knowledge in remote sensing studies, unsupervised
classification and cluster analysis have great importance compared to supervised
classification \cite{Abbas2016}.

Unsupervised learning is a machine learning technique used when the learning
algorithm has no pre-existing labels and looks for patterns in the data set with
minimum human supervision and without external guidance. In contrast to
supervised learning, where a training set of inputs and observations are given
to the method, unsupervised learning uses only the inputs and should find an
interesting pattern in the data set. Two main approaches are used in
unsupervised learning, named principal component \cite{Mallapragada2018} and
cluster analysis \cite{Xu2015}. In the clustering method, unlabeled and unsorted
data points are grouped and segmented according to similarities and differences
to find structure in the inputs, group it into meaningful clusters, and
understand the data set better. There are different algorithms for Clustering,
such as K-Means, fuzzy C-Means, ISODATA \cite{Ball1965}, and so on. In the
K-Means algorithm, the goal is to minimize the sum of the squared distances
points have in the data set from its center. The problem with the K-Means
algorithm is that it requires a priori knowledge about the number of classes in
the data set, and choosing the cluster centers randomly may not result in
optimal partitioning of the data set.

One of the most popular clustering algorithms is called ISODATA clustering,
based on the K-means algorithm with some variations. The ISODATA method's
preference in comparison with the K-means algorithm is that the ISODATA
algorithm uses splitting and merging forms for clusters. The ISODATA method
allows starting from arbitrary cluster centers and obtains better partitioning
results than the K-Means algorithm, but it is not entirely resilient to initial
cluster centers. In this paper, a novel method using the ISODATA approach of
splitting clusters to select quality initial cluster centers is proposed.

% The paper is organized as follows: Section 2 introduces K-Means and ISODATA
% clustering algorithms. The new proposed algorithm is explained in Section 3. We
% present the experimental results in Section 4. Finally, Section 5 contains
% concluding remarks.

@@

This method began as a modified ISODATA process. However, the modification was
unique in that there is a clear distinction from the proposed change and the
beginning of the original process. Although more complex than other methods of
generating initial cluster centers, this approach guarantees variance between
clusters. As a result, ISODS can be used to generate clusters for other
clustering algorithms, such as K-Means.

@@

Due to the response of the ISODATA algorithm to the discarding, merging, and
splitting parameters, these parameters should be carefully selected. Since the
goal of this research is to explore the impact of initial cluster centers on
ISODATA, we use a relatively simple approach for generating ISODATA parameters.
Based on the available training data with included labels, each class is
summarized with the mean and standard deviation of that class, as well as the
number of members in each class. Typically, since we are working with
unsupervised learning algorithms, class information may not be available.

To determine $\theta n$, the minimum number of cluster members used for
discarding, we find the class with the fewest members and use the number of
members in that class as $\theta n$. For $\theta e$, the maximum standard
deviation allowed for a cluster, we follow a similar approach and find the
maximum standard deviation within a class. Finally, for $\theta c$, we calculate
the distance between the means of each class, and use the minimum distance as
$\theta c$.

For experiments 1 and 2, we explore applying ISODS to ISODATA and K-Means by
using benchmark UCI datasets Iris, Wine, and Glass. We also show results from
applying Random Centers and K-Means++ Centers. To evaluate the cluster results,
external cluster validation measures are used. We utilize pair-counting measures
which rely on counting matching pairs between two sets of data to determine
agreement and disagreement \cite{Rezaei2016}. The pair-counting measures applied
are \textit{rand index} (RI - agreement probability), \textit{adjusted rand
  index} (ARI), \textit{mirkin's index} (MI - disagreement probability), and
\textit{hubert's index} (HI - $P(agree) - P(disagree)$)\cite{Hubert1985}
\cite{Meila2003}.

@@

@@

\bibliographystyle{unsrt}
\bibliography{paper}
